% appendix_continuity.tex
% Core v2.5 — Appendix B: Continuity and Transition Operators

\section{Appendix B: Continuity and Transition Operators}
\label{sec:appendix-continuity}

This appendix provides the expanded formal structure underlying continuity,
dimension growth, metaspace embedding and collapse across the hierarchy
\(
K_0 \to K_1 \to \dots \to K_{12}.
\)

\subsection{Transition Operators}
A transition between consecutive continua is defined by
\[
  \Psi_{x\rightarrow x+1} : K_x \longrightarrow K_{x+1},
\]
acting on all structural components:
\[
  \Psi_{x\rightarrow x+1}(\Omega_x, A_x, P_x, \Theta_x, J_x, C_x, k_x)
  =
  (\Omega_{x+1}, A_{x+1}, P_{x+1}, \Theta_{x+1}, J_{x+1}, C_{x+1}, k_{x+1}).
\]

\subsection{Continuity Conditions}

A transition is \emph{continuous} if the five structural conditions (C1)–(C5)
hold:

\paragraph{(C1) Admissibility preservation.}
\[
  \Omega_x \subseteq \Omega_{x+1}.
\]

\paragraph{(C2) Axis extension.}
\[
  A_x \subset A_{x+1}.
\]

\paragraph{(C3) Potential inheritance.}
\[
  P_x \subset P_{x+1}.
\]

\paragraph{(C4) Threshold inheritance.}
\[
  \Theta_x \subseteq \Theta_{x+1}.
\]

\paragraph{(C5) Cycle embedding.}
\[
  C_x \subseteq C_{x+1}.
\]

Under these conditions,
\[
  k_x(t)>0 \;\Longrightarrow\; k_{x+1}(t)>0.
\]

\subsection{Dimension Emergence}

A new structural dimension appears at transition \(K_x\to K_{x+1}\) only if the
following criteria are jointly satisfied:

\paragraph{(D1) Structural incompatibility.}
Existing axes cannot encode a new class of differences.

\paragraph{(D2) Dimensional tension threshold.}
\[
  T_x(s) \ge \Theta_x^{\mathrm{dim}}
  \quad\text{for some } s\in\Omega_x.
\]

\paragraph{(D3) Metaspace admissibility.}
Both continua must embed into the next metaspace:
\[
  \Omega_x \subseteq \Omega(M_{x+1}),
  \qquad
  \Omega_{x+1} \subseteq \Omega(M_{x+1}).
\]

If (D1)–(D3) hold, then
\[
  \dim K_{x+1} = \dim K_x + 1.
\]
This enforces monotonic dimension growth (Theorem: Monotonicity of Dimension).

\subsection{Collapse Conditions}

A continuum collapses at time \(t^\ast\) if and only if at least one of the
following holds:

\paragraph{(CL1) Loss of admissibility.}
\[
  \Omega_x(t^\ast)=\varnothing.
\]

\paragraph{(CL2) Loss of cycles.}
\[
  C_x(t^\ast)=\varnothing.
\]

\paragraph{(CL3) Global threshold violation.}
\[
  T_x(s)>\Theta_x(s)
  \quad\text{for all } s\in S(K_x).
\]

\paragraph{(CL4) Metaspace violation.}
\[
  \Omega_x \not\subseteq \Omega(M_x).
\]

Collapse implies:
\[
  k_x(t^\ast)=0.
\]

\subsection{Time Continuity}

Let the intrinsic cycle periods of \(K_x\) be \(\{\tau_n^{(x)}\}\).  
Time remains continuous across transitions if
\[
  \tau(K_{x+1}) \le \tau(K_x)
  \quad\text{or}\quad
  \tau(K_{x+1}) = \tau(K_x).
\]

If the characteristic cycle duration diverges,
\[
  \tau(K_x)\to +\infty,
\]
the continuum is nonviable (collapse through loss of cycles).

\subsection{Path Continuity in State Space}

A trajectory \(s(t)\) is continuous across levels if
\[
  s(t)\in\Omega_x
  \quad\text{and}\quad
  \Psi_{x\rightarrow x+1}(s(t))\in\Omega_{x+1},
\]
for all relevant times.

\subsection{Metaspace Chain Conditions}

Each continuum embeds into its metaspace:
\[
  \Omega_x \subseteq \Omega(M_x), 
  \qquad 
  A_x \subseteq A(M_x).
\]

Moreover, the degrees of freedom satisfy:
\[
  \mathrm{DoF}(K_x)\le \mathrm{DoF}(M_{x+1}).
\]

\subsection{Compact Summary}

Transition operators generate the chain:
\[
  K_0 \xrightarrow{\Psi_{01}} K_1
      \xrightarrow{\Psi_{12}} K_2
      \dots
      \xrightarrow{\Psi_{11,12}} K_{12}.
\]

Continuity requires:
\[
  \Omega_x\subseteq\Omega_{x+1},\ 
  A_x\subseteq A_{x+1},\
  P_x\subseteq P_{x+1},\
  \Theta_x\subseteq\Theta_{x+1},\
  C_x\subseteq C_{x+1},
\]
and ensures:
\[
  k_x>0 \Rightarrow k_{x+1}>0.
\]

Dimension is monotonically increasing, and collapse occurs only through loss of
admissibility, violation of all thresholds, loss of cycles or metaspace
incompatibility.
